---
title: "Listening to the price-setters: Inferring inflation expectations from synthetic surveys"
author: "Daniel Aromí^[IIEP UBA-Conicet, Facultad de Ciencias Económicas, Univ. de
Buenos Aires. Facultad de Ciencias Económicas, UCA, Buenos Aires, Argentina. aromi.daniel@gmail.com], Paula Bonel^[IIEP UBA-Conicet, Facultad de Ciencias Económicas, Univ. de Buenos Aires. Facultad de Ciencias Económicas, UCA, Buenos Aires, Argentina. mpaulabonel@gmail.com], Llada Martin^[IIEP UBA-Conicet, Facultad de Ciencias Económicas, Univ. de Buenos Aires. lladamartin@gmail.com]"
date:  "November, 2024"
output:
  #pdf_document:
  #  toc: yes
   # toc_depth: '3'
  html_document:
    theme: united
    #toc: yes
    #toc_depth: 3
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "C:/Users/user/Documents/GitHub/ECC")
```

## Inflation expectations indicator based on synthetic surveys
Inflation dynamics are driven by the way in which a diverse set of economic agents
perceive the economic environment. Estimations of these perceptions are key features in academic and policy analyses and, as a consequence, important resources have been devoted to improving these approximations.

Prominent indicators of expected inflation are given by professional surveys, business surveys and market-based expectations. These indicators are shown to
provide useful information (Berge, 2018; Candia, Coibion, & Gorodnichenko, 2023,
2024; de Mendon¸ca, Garcia, & Vicente, 2021).

In this website, we propose a novel method to extract business inflation expectations via synthetic surveys completed by finetuned large language models (LLMs). The document could be download from [here](https://papers.ssrn.com/sol3/papers.cfm?abstract_id=4955056).

Inflation has once again become a major concern for economic agents. The following table shows the latest numbers:

```{r code1,include=F, warning=FALSE }
#load libraries
library(tidyverse)
library(gt)
library(gtExtras)
library(plotly)
library(ggrepel)
```

```{r code2,include=FALSE, warning=FALSE }
# base surge de script "script_reg_plan.R" en ".\reuters\taller paper\ejercicio"
df = read.csv("./docs/dataset_report.csv")
```


```{r code3,include=F, warning=FALSE,echo=FALSE}
#table gt 

last_date_cpi <- df %>% 
   slice(n()) %>% 
  mutate(gcpi_annual = gcpi_annual*100) %>%
  mutate(country='US',Last = gcpi_annual) %>%
  select(country,Last)

last_period_cpi <- df %>% 
  filter(substr(date,1,4)=="2023") %>% 
  mutate(gcpi_annual = gcpi_annual*100) %>%
  summarise(Timeline = list(c(gcpi_annual)))%>%
  mutate(country='US') %>%
  select(country,Timeline)


tabla_cpi = last_date_cpi  %>% 
              left_join(last_period_cpi, by = 'country') %>% 
              gt() %>% 
              tab_header(
                title = 'Annual inflation rate in the US'
              ) %>% 
               tab_caption("Monthly report") %>%
              tab_footnote(footnote = md("Source: [St. Louis Fed](https://fred.stlouisfed.org/series/CPIAUCSL)")) %>%
              cols_label(Last = 'Inflation rate (last)',
                         Timeline = 'Period') %>%
              fmt_number(columns = where(is.numeric), decimals = 2) %>%
              gt_theme_538() %>% 
              gt_plt_sparkline(
                column = Timeline,
                palette = c("grey40", "grey40", "grey40", "dodgerblue1", "grey40"),
                fig_dim = c(5, 28)
              ) %>%
              cols_hide('country')


```
```{r code4,include=T, warning=FALSE,echo=FALSE}
tabla_cpi
```

In the following figure, we show a strong association between inflation and synthetic business inflation expectations. A particularly noticeable feature is the increase in both indicators observed in 2021, which is later reversed by the end of the sample period. Consistently, the correlation coefficient between the time series is 0.55.

```{r code5,include=F, warning=FALSE,echo=FALSE}

#agregar label al grafico
df <- df %>%
  mutate(mylabel = paste("Date: ", date, "\n",
                         "Monthly inflation rate: ", round(gcpi,2), "\n",
                         "SBS index: ", round(ss3,2),sep=""))


df = df %>% mutate(
    ss3_scaled = scale(ss3)[, 1],
    gcpi_scaled = scale(gcpi)[, 1]  
  )

df = as_tibble(df)

df2 = df %>% select(date,mylabel, ss3_scaled, gcpi_scaled)
df2 = df2 %>% gather(var,value,3:length(df2))


#labels = c("gcpi_scaled" = "Inflation Rate", "ss3_scaled" = "SBS Index")
df2$var2 = ""
df2$var2[df2$var=="gcpi_scaled"] = "Inflation Rate"
df2$var2[df2$var=="ss3_scaled"] = "SBS Index"


# Crear el gráfico con ggplot y usar "text" en lugar de "label"
p <- ggplot(df2) +
  geom_line(aes(x = as.Date(date), y = value, color = var2)) +
  geom_point(aes(x = as.Date(date), y = value), alpha = 0) + # Añadir puntos invisibles para que plotly pueda leer los 'hover'
scale_x_date(date_labels = "%Y-%m-%d",date_breaks  ="6 month")+ 
  labs(x = "",y="", title = "",color = "") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
axis.line = element_line(colour = "black"),
panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank())+
  scale_color_brewer(palette = "Dark2")  # Asignar nuevos nombres a las categorías


# Convertir a ggplotly y pasar directamente los valores del tooltip
p_plotly <- ggplotly(p) %>%
  layout(
    hoverlabel = list(bgcolor = "white", font = list(size = 12)),
  annotations = list(
      x = 0.15,  # Centramos la nota al pie
      y = -0.22,  # Un poco más cerca del gráfico
      text = "Note: the variables were standardized.",
      xref = 'paper',
      yref = 'paper',
      showarrow = FALSE,
      xanchor = 'center',  # Centramos horizontalmente la nota al pie
      yanchor = 'top',
      font = list(size = 10)
    )
  ) %>%
  style(text = df2$mylabel, hoverinfo = "text")


```


```{r code6,include=T, warning=FALSE,echo=FALSE}
p_plotly
```

An Paula additional motivation for this project is to provide the community with the set of created indicators (downloadable [here]()). This database will be updated regularly. In the future, we also hope to provide the LLMs so that they can be used by the community.


## References
Berge, T.J. (2018). Understanding survey-based inflation expectations. International Journal of Forecasting, 34 (4), 788–801.

Candia, B., Coibion, O., Gorodnichenko, Y. (2023). The macroeconomic expectations
of firms. In handbook of economic expectations (p. 321-353). Academic Press.

de Mendonça, H.F., Garcia, P.M., Vicente, J.V.M. (2021). Rationality and anchoring
of inflation expectations: An assessment from survey-based and market-based
measures. Journal of Forecasting, 40 (6), 1027–1053.